An IFS study of Low-powered Radio Galaxies

### 2018-01-21 paper.tex
EoDC: Introduction and maps
 On branch master
 Changes to be committed:
	modified:   paper.tex
	modified:   produce_maps.py

	We refer to the combined Atlas$^\text{3D}$ and MASSIVE sample as the A+M sample.

>>>>

	It is becoming increasingly clear that active galactic nuclei (AGN) exist in two different modes, radiative and jet mode \citep[e.g.][]{Antonucci2012, Heckman2014}, raising speculation about differing mechanisms and possibly fuel sources for each. The former is thought to be the powerful engine that initially and rapidly shuts off star formation \citep[e.g.][]{Thomas2005, Thomas2010}, either by removing the cold gas from the galaxy entirely, heating it such that it joins the hot X-ray gas often present in massive galaxies \citep[e.g.][]{OSullivan2001}, or rendering it unusable for star formation (such as by stabilising it against collapse; e.g.\ \citealt{Martig2009}). After a single radiative-mode episode, the AGN is thought to subside to become jet dominated, repeatedly reheating the surrounding ISM to counter any cooling that might allow the cold gas reservoir to replenish (and subsequently form stars).

	There is a strong consensus within the scientific literature that the most powerful radio galaxies are the result of gas-rich, major mergers causing large amounts of material to fall on to the black hole at the centre of the resulting galaxy, thus producing an immensely powerful AGN include a powerful radio jet \citep[e.g.][]{Malin1983, Quillen1992, Lim2000}. However, such galaxies are by construction extremely rare and are not representative of the general radio galaxy population. Most radio galaxies have low power, and their basic properties and origins are poorly constrained. One of the most popular suggestions is that they are bona-fide detections of jet-mode AGN.%, a class of AGN thought to maintain the dearth of star formation in early-type galaxies (ETGs), after most of the fuel for star formation has been removed by an earlier event. 
	They are known to be dusty with dust lanes having a tendency to be perpendicular to the jets \citep[e.g.][]{DeRuiter2002, VerdoesKleijn2005}. They have little ionized gas in the interstellar medium (ISM) \citep[e.g.][]{}, but often have significant quantities of hot X-ray gas \citep[e.g.][]{} (the cold phase of the ISM of radio galaxies will be examined in Ruffa et.al.\ in prep.). If the low star-formation rates typically seen in massive early-type galaxies (ETGs) are preserved by such jet-mode AGN, then all such ETGs must regularly go through similar active phases, and therefore low-powered radio galaxies should be much the same as their radio-quiet (quiescent phase) ETG counterparts.

	% In jet-mode (also called radio- or maintenance-mode in the literature) AGN, the dominant form of emitted energy away from the black hole is the kinetic (i.e.\ mechanical) energy of the material that makes up the jet. Jet-mode AGN can be directly observed `feeding back' to their host galaxies, as jets often terminate at shocks (hot spots) that make up the walls of cavities observed in the hot (X-ray) haloes of the host galaxies.

	% Jet-mode AGN tend to be radiatively inefficient, with Eddington ratios $<$ 1\%. This implies that the innermost regions of the accretion discs must be geometrically thick, with inflow times much less than the radiative cooling time-scales. The resulting inflows are known as advection-dominated accretion flows (ADAF; \citealt{Narayan1994}). Such systems are capable of launching jets. It is possible that there is a thin disc beyond the thick disc (further from the black hole), but it must be truncated by the thick disc \cite[e.g.][]{Abramowicz2002, Sadler2014}. Much of this is very difficult to resolve observationally, as the spatial scale is $< 100$ Schwarzschild radii, and many of the models remain degenerate \citep[e.g.][]{Quataert1999}.

	% There is still much speculation about the nature of the fuel (and its origin) in jet-mode AGN, though a consensus seems to building in favour of an internal origin. Many people advocate hot gas as being the ultimate original fuel \citep[e.g.][]{Allen2006}. \citet{Hardcastle2007} suggest that it could be directly accreted through Bondi accretion \citep{Bondi1952}, a simple model where gas is accreted in a spherically-symmetric manner directly from the hot phase on to the black hole. Others invoke a cooling mechanism to allow the X-ray gas to condense into molecular (i.e.\ cold) gas before accretion. Yet others point to a higher higher detection rate of dust and molecular gas in RGs compared to the general population \citep[e.g.][]{Lim2000, DeRuiter2002, Leon2003, VerdoesKleijn2005}, suggesting that cold gas from either internal reservoirs or externally-accreted may be the fuel. \citet{DeRuiter2002} showed that, when dust is present, its mass is correlated with the radio power and there seems to be a bias towards dust lanes orientated orthogonally to radio jets \citep{VerdoesKleijn2005}.

	% Beyond this, the source of the energy for powering the jet is still unclear. Energy may be extracted from the black hole via the \citet{Blandford1977} mechanism or variations thereof \citep[e.g.][]{Koide2002}, or from the accretion disc as discussed by \citet{Blandford1982} and others \citep[e.g.][]{Hujeirat2003}.

	% It is thought that the duty cycle of a jet is very short; the jet is turned on for very short periods of time before turning off for extended periods.

	% An obvious feedback mode is seen in the haloes of hot X-ray gas around galaxies, where radio jets terminate. The kinetic impacts of the jets create cavities or bubbles in the hot gas (first conceptualised by \citealt{Gull1973}). In cluster environments, these are buoyant and can separate, rising through the intra-cluster medium \citep[e.g.][]{Churazov2000, Churazov2001, McNamara2000}. However, the bubbles are intrinsically anisotropic, while star formation must be suppressed on a global scale. They also exhibit steep abundance gradients, showing that there is little mixing that could disperse energy via convection. Instead, from observations of the very brightest X-ray clusters such as Perseus \citep{Fabian2003, Fabian2006}, Virgo \citep{Forman2007}, Centaurus \citep{Sanders2008}, Abell 2052 \citep{Blanton2011} and simulations \citep[e.g.][]{Ruszkowski2004, Sijacki2006}, it appears that sound waves in the ISM are generated by the repetitive blowing of bubbles. The energy of the waves is likely sufficient to offset that lost by the gas via normal cooling mechanisms \citep{Fabian2003}. The low amplitude of these waves means that they are not detectable in all but the closest clusters (more distant clusters would require in excess of 10 days of continuous \textit{Chandra} observations; \citealt{Graham2008}).

	% Even when the jets are turned off or only very weak, the galaxies may still contain low ionization nuclear emission-line regions (LINERs). It is worth noting that while the most powerful LINERs are likely to be low-luminosity AGN, many (indeed most SDSS-classified LINERs) are not; they belong to a recently-defined category known as LIERs i.e.\ the emission-line region is not nuclear (LINERs but without the `N'; e.g.\ \citealt{Sarzi2005, Sarzi2010, Singh2013, Belfiore2016a}). Such emission is mostly attributed to ionization from radiation from post-asymptotic giant branch (pAGB) stars. \citet{Capetti2011} suggested using an equivalent width of the [\ion{O}{iii}] emission line of $EW_\text{[\ion{O}{iii}]} = 1$\,\AA\ as a threshold for LIERs; galaxies with greater equivalent widths are likely low-luminosity AGN.

	% As stated above, most radio emission from RGs results from synchrotron emission from the relativistic plasma within the jets, moving around the magnetic fields tangled in the jets. This means that all such RGs can be considered a jet detection (ignoring the very low radio-luminosity galaxies resulting from high levels of circumnuclear star formation). RGs are always jet-mode AGN, except at the highest radio luminosities where the bright radio jets are dominated (in terms of energy output) by the brighter still accretion disc. These are radiative-mode AGN (see Section \ref{subsubsec:RadiativeRadio}).

	% All radio-selected AGN are massive, with very old stellar populations and high concentration indices \citep[e.g.][]{Best2005}. \citet{Best2005} also found that the fraction of galaxies with detectable radio emission is related to the galaxies' stellar and black hole masses by power laws with indices of 2.5 and 1.6, respectively.

	To test this hypothesis, a complete sample of 11 nearby low-powered radio galaxies was constructed and observed with a suite of instruments. Here, we present the spatially-resolved stellar kinematics, stellar populations, and ionized gas distribution, kinematics and ionization derived from integral-field spectroscopic (IFS) observations with the Visible Multi-object Spectrograph (VIMOS) and archival data from the Multi-unit Spectroscopic Explorer (MUSE), both on the Very Large Telescope (VLT). We compare the characteristics and properties (accessible with $V$-band IFS data) of our sample of low-power radio galaxies to those of ordinary (i.e.\ optically-selected) ETGs from the Atlas$^\text{3D}$ and MASSIVE IFS surveys. We hereafter refer to the combined Atlas$^\text{3D}$ and MASSIVE sample as the A+M sample.

	% For the stellar component of our sample galaxies, we use the kinematic maps to identify regular/non-regular rotators and kinematic substructures. In our sample 4 galaxies are regular rotators and two (with a tentative third) galaxies were found to contain kinematically decoupled cores (KDCs). We use the $\lambda_R$ specific angular momentum parameter from Atlas$^\text{3D}$ to classify the galaxies according to the fast/slow rotator classification scheme. We found a similar fraction ($45\pm13$\%) of slow rotators to the combined Atlas$^\text{3D}$ and MASSIVE samples, corrected for mass distribution. We also find the fast rotators tend to have aligned kinematic and photometric position angles, while the slow rotators can have any angle between the two position angles. The best-fitting stellar populations show our sample to be typical massive ETGs: i.e.\ mostly old, metal rich and alpha-element enhanced. We found and measure the radial stellar population gradients and found averages of $\Delta_\text{log age} = 0.014\pm0.007 \,\mathrm{dex \, arcsec^{-1}}$ and $\Delta_\text{[Fe/H]} = -0.03\pm0.01 \, \mathrm{dex \, arcsec^{-1}}$. The KDCs found previously were observed to be large and contain old stellar populations and are the result of galaxy mergers.

	% Some of the literature suggestion that if cold gas is the fuel for the AGN then unless it is accreting in a highly turbulent way, it might be expected to form stars. However we see no evidence of a very young stellar population right at the centre of our sample galaxies.

	% For the ionized gas components of our sample galaxies, the estimated gas masses were found to be in the range of $10^4$\,--\, $10^6\,\mathrm{M_\odot}$, which is the upper end of the range expected from optically-selected ETGs. This was the only suggestion we find that our sample galaxies may be intrinsically different than optically-selected ETGs. We detected spatially extended gas in 4 of our sample, with a range of kinematics. Two galaxies had ionized gas kinematics misaligned to that of the stars, suggesting and external origin to the gas, while the kinematics of one was aligned with that of its stars (consistent with an internal origin). The final galaxy was unsettled and thus not clear.

	% Using the classic Baldwin\,--\,Phillips\,--\,Terlevich (BPT) diagnostic plots, as well as similar alternatives, for galaxies with observations outside of the wavelength range required for the BPT plots, we find the dominant ionization source for each galaxy. We find that 9 of our sample galaxies are low-ionized nuclear emission region galaxies (LINERs; 5 of which can be attributed to the radiation field from the AGN), 1 is a Seyfert II and 1 a passive (line-less) galaxy.

	% All of the properties found are consistent with that of optically selected ETGs with a similar mass distribution. We thus find no conclusive difference between the stellar and ionized gas properties all of these ETGs except for a tentative suggestion of slightly higher ionized gas masses in our radio galaxies. If confirmed, the latter may be gas built up during the quiescent phase of ETGs, ultimately triggering the active (radio) phase itself then re-removing such gas from the galaxy.

	% Some of the sample galaxies have clear evidence of merger events (e.g.\ large KDCs), but others have a complete lack of evidence of a merger, suggesting that radio galaxies can form without the requirement of a merger event. Indeed, it is well known that NGC 612 harbours a large extended gas disc with kinematics that are consistent with an internal origin of the gas. As such, in rare cases, it is possible that radio galaxies can be fueled through entirely secular processes.
